### 2020-05-17 Papers/01_interf_transf/Interference Transform.tex
started paper 01 presenting the interference transform
added various explorations into a closed form estimation of max frequency in the folder python/explorations/cf

\documentclass{article}

\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}

\title{Interference Transform: Estimating the frequency and phase of low resolution samples}
\author{Carlos Tarjano}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

{\bf Keywords:}

\section{Introduction}

\section{Bibliographic Review}

\subsection{Sound Representation}
frequency domain, one hot encoding, spectrogram

\subsection{Computer Graphics}
Bezier, splines, some raster, image rotation

\subsection{Sound synthesis}

\section{Methodology}

\section{Results}

\section{Discussion + Conclusion}

\section{References}

\end{document}