### 2020-10-07 Text/00_Include.tex
organized text division

\documentclass[12pt]{article}
\usepackage[a4paper, top=3cm, left=3cm, bottom=2cm, right=2cm]{geometry}
\usepackage[utf8]{inputenc}
% \usepackage{fontspec}
% \setmainfont{Arial}
\title{Sound Synthesis}
\author{Carlos Tarjano}