### 2020-10-08 Text/00_Include.tex
structured thesis text in separated .tex files

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[backend=biber]{biblatex}

\usepackage[a4paper, top=3cm, left=3cm, bottom=2cm, right=2cm]{geometry}
\usepackage{indentfirst}
\setlength{\parindent}{1.25cm}

\usepackage[no-math]{fontspec}
\defaultfontfeatures{Ligatures=TeX}
\setmainfont{Tex Gyre Termes}

\usepackage[nodisplayskipstretch]{setspace}
\setstretch{1.427465}

\usepackage[explicit]{titlesec}
\titleformat{\section}{\normalfont\bfseries}{\thesection}{1em}{\MakeUppercase{#1}}
\titleformat{\subsection}{\normalfont}{\thesubsection}{1em}{\MakeUppercase{#1}}
\titleformat{\subsubsection}{\normalfont\bfseries}{\thesection}{1em}

\usepackage{amsmath}  % enables declaremathoperator
\DeclareMathOperator{\sign}{sign}

\usepackage{amsfonts} % \mathbb
\usepackage{graphicx} % manipulate imported images

\usepackage{hyperref}
\usepackage{hypcap}
\hypersetup{
  colorlinks=true, 
  allcolors=black,
  pdfauthor={Carlos Henrique Tarjano Santos},
  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
}

\bibliography{00_BIBLI.bib}

\title{Sound Synthesis}
\author{Carlos Tarjano}

% \usepackage[T1]{fontenc}
% \usepackage{authblk}