### 2020-10-09 Text/00_Include.tex
English set-up ok

\documentclass[12pt]{article}

% \usepackage[utf8]{inputenc}
% \usepackage[T1]{fontenc}
% \usepackage{authblk}

\usepackage[backend=biber]{biblatex}
%\usepackage{biblatex}

\usepackage[a4paper, top=3cm, left=3cm, bottom=2cm, right=2cm]{geometry}
\usepackage{indentfirst}
\setlength{\parindent}{1.25cm}

\usepackage[no-math]{fontspec}
\defaultfontfeatures{Ligatures=TeX}
\setmainfont{Tex Gyre Termes}

\usepackage[nodisplayskipstretch]{setspace}
\setstretch{1.427465}

\usepackage[explicit]{titlesec}
\titleformat{\section}[hang]{\normalfont\bfseries}{\thesection}{1em}{\MakeUppercase{#1}}
\titleformat{\subsection}[hang]{\normalfont}{\thesubsection}{1em}{\MakeUppercase{#1}}
\titleformat{\subsubsection}[hang]{\normalfont\bfseries}{\thesubsubsection}{1em}{#1}%{\MakeLowercase{#1}}

\usepackage{titletoc}
\titlecontents{section}[1em]{\vspace{12pt}}{\bfseries\normalsize\contentslabel[\thecontentslabel]{3em}\uppercase}{\hspace*{-2em}\uppercase}{\titlerule*[.75em]{.}\contentspage}
\titlecontents{subsection}[1em]{\vspace{12pt}}{\normalfont\normalsize\contentslabel[\thecontentslabel]{3em}\uppercase}{\hspace*{-2em}\uppercase}{\titlerule*[.75em]{.}\contentspage}
\titlecontents{subsubsection}[1em]{\vspace{12pt}}{\bfseries\normalsize\contentslabel[\thecontentslabel]{3em}}{\hspace*{-2em}}{\titlerule*[.75em]{.}\contentspage}

\usepackage{amsmath}  % enables declaremathoperator
\DeclareMathOperator{\sign}{sign}

\usepackage{amsfonts} % \mathbb
\usepackage{graphicx} % manipulate imported images

\usepackage{fancyhdr} % \pagestyle{fancy}
\fancyhf{}
\fancyhead[R]{\thepage}
\renewcommand{\headrulewidth}{0pt}

\usepackage[toc,section=section]{glossaries}
\makeglossaries

\usepackage{hyperref}
\usepackage{hypcap}
\hypersetup{
  colorlinks=true, 
  allcolors=black,
  pdfauthor={Carlos Henrique Tarjano Santos},
  pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
}

\bibliography{00_BIBLI}

\title{Sound Synthesis}
\author{Carlos Tarjano}

% \newglossaryentry{VST}{name=VST, description={Virtual Studio Technology}}