### 2021-02-25 Papers/02_Sound_Compression/SoundRepresentation.tex
some progress on the paper

\documentclass{article}
\usepackage{lmodern} % Latin Modern Font
\usepackage[T1]{fontenc}
\usepackage{amsfonts} % \mathbb
\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
% \graphicspath{{images/}}
\usepackage[backend=biber]{biblatex}
\usepackage{amsmath}
\usepackage[makeroom]{cancel}
% \usepackage{auto-pst-pdf}

% \usepackage{hyperref}
% \usepackage{hypcap}
% \DeclareMathOperator{\sign}{sign}
% \hypersetup{
%   colorlinks=true, 
%   allcolors=black,
%   pdfauthor={Carlos Henrique Tarjano Santos},
%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
% }

\bibliography{bibli}

\usepackage{authblk}
\title{Title}
\author[1]{Carlos Henrique Tarjano Santos}
\author[2]{Valdecy Pereira}
\affil[1]{(corresponding author, carlostarjano@id.uff.br) Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
\affil[2]{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

\begin{document}

\maketitle

\begin{abstract} % ≅ 250 words  
% Clear and simple miniature version of the paper
% Self-contained
% Past tense for the current study
% Present tense to refer to stablished facts

% Background (What is already known about the subject)
Sample based virtual instruments are the de facto standard for the emulation of real word acoustic instruments. Nevertheless little attention is given to the encoding of the samples used in those instruments; in fact, lossless containers like wav or proprietary formats are common, pushing the on disk size of those libraries to sizes exceeding the gigabytes.
% Motivation (Why do we care about the problem and the results?) 
This renders the most realistic digital instruments inaccessible in many but the most high-end hardware; controllers, like electronic drum kits and controller keyboards and even less robust personal computers might encounter problems implementing those software.
% Problem | Objectives (What problem is the paper trying to solve?) (Scope)
We propose a lossy compression algorithm tailored specifically to highly uniform sounds, as those present in these instruments, due to the nature of the sounds, highly redundant, substantial improvements can be obtained, with minimal quality loss.

% Methods | Procedure | Approach (What did you actually do to get your results?)

% Results | Findings | Product (what did you learn | invent | create?)

% Conclusions | Implications (What are the larger implications of your findings)
The method, being physically informed, also serves as base 
lends itself well to machine learning, as the sound can be represented by an evolving waveform over time (pseudo-cycles)

\end{abstract}

{\bf Keywords:}

\section{Introduction} % ≅ 500 words
% ≅ 10\% of the text
% broad to specific
% show that the research question is clear, concise, and worthy of study.
% allow the reader to understand the rest of the paper without referring to previous publications on the topic
% What did you do?
% conclusions

% Opener sentence (Question | Assertion | Intriguing Revelations)
Many of the proposed benefits of an electronic drumkit, like portability and practicity, are compromised when, to obtain a realistic sound on par with the acoustic counterpart, the output of the drumkit has to be processed in another piece of hardware, such as a personal computer.

The same is true for digital pianos.

This transforms the instrument in a mere controller, robbing it of characteristics that provide the personality of the conventional instruments, such as a particular timbre.

% Literature review | Brief Context of Prior Research
  % put your research in the context of other research
  % Nature and scope of the problem investigated
  % give them a framework for understanding it
  % What is the problem to be solved?
  % Are there any existing solutions?
  % Which is the best?
  % What is its main limitation?
  % What do you hope to achieve?

% Restate Your Question as Something Not Known or Fully Understood by Prior Research
  % identify the gap in the literature
  % use but, however, etc…

% State the Significance of Your Question
  % make readers understand why they should read your report

% State Your Claim | Objective | Hypothesis
  % should clearly relate to the information gap

% brief outlook on the structure of the paper

\section{Methodology} % ≅ 1000 words
  % How did you do it?
  % Justify choices
  % Describe the work so as people can reproduce it

% Overview of methods
We can interpret a continuous or discrete signal, as long as it is at least pseudo periodic, as succession of waveforms, each one representing a (pseudo) cycle of the wave.

We start with an estimate of the waveform of the signal being processed.

% Restate the purpose of the work

% Essential background information

% Relate to other studies

% Indicate where problems occurred

\section{Results} % ≅ 3000 words
  % What did you find?
  % Use figures, tables, etc.
  % Clearly and simply stated.

% Revisit purpose and methods

% Overview of results

% Key results in details

% Comparison with results in other works

% Comparison with predictions

% Problems with results

% Possible implications

\section{Discussion} % ≅ 2000 words

% Put results into a broader context.

% specific to broad

% What does it all mean?

% an interpretation of how each of the experiments supported your central hypothesis

% add in supporting findings from the literature

% what principles have been established or reinforced

\section{Conclusion} % ≅ 250 words
  % leave readers with a clear idea of your claim
  % make readers understand its importance
  % suggest further research
  % how the work advances the field?
  % significance and implications of the results
  % implications for broader application
  % research opportunities

% Restate Your Claim

% Point Out: New Significance | Practical Application | New Research

% \begin{figure}[ht!]
%   \centering
%      \includegraphics{images/01signalenvelope.pdf}
%   \caption{A discrete wave of the singing voice of an alto singer. The black lines are the frontiers; in the detail view, it can be seen that each region between two diamonds comprises a pseudo-cycle, as defined by the points belonging to the upper frontier. Similarly, two adjacent squares delimitate a pseudo-cycle, from the standpoint of the lower frontier.}
%   \label{fig:signalenvelope}
% \end{figure}

\printbibliography % 20 ⇔ 60 itens

\end{document}