### 2021-03-14 THESIS.tex
initial

\input{00_Include.tex}

\begin{document}

%\include{01_Resumo.tex}
\pagestyle{empty}
\include{02_Abstract.tex}

% \newpage
% \listoffigures % lista de ilustracoes

% \newpage
% \listoftables % lista de tabelas

\newpage
\tableofcontents

\newpage
\pagestyle{fancy}
\include{03_Introduction.tex}
\include{04_Theory.tex}
\include{05_Characterization.tex}
\include{06_SignalEnvelope.tex}

% \section{References}
\printbibliography %[heading=none]

\newpage
\printglossaries

\end{document}