### 2021-05-03 Papers/01_Pulses_Envelope/PulsesEnvelope.tex
progress with the revision

\documentclass{article}
\usepackage{float}
\usepackage{lmodern} % Latin Modern Font
\usepackage[T1]{fontenc}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage{amsfonts} % \mathbb
\usepackage{amssymb} % \therefore
\usepackage[margin=0.5in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[backend=biber, sorting=none]{biblatex}

\usepackage{algorithm}
\usepackage{algorithmic}

% \usepackage{hyperref}
% \usepackage{hypcap}
% \hypersetup{
%   colorlinks=true, 
%   allcolors=black,
%   pdfauthor={Carlos Henrique Tarjano Santos},
%   pdftitle={Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Real Signal}
% }

\DeclareMathOperator{\sign}{sign}
\bibliography{bibli}

\title{Envelope Estimation using Geometric Properties of a Discrete Real Signal}
\author[1]{Carlos Henrique Tarjano Santos}
\author[2]{Valdecy Pereira}
\affil[1]{(corresponding author, carlostarjano@id.uff.br) Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
\affil[2]{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

\begin{document}

\maketitle

\begin{abstract}
% Background
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example.  
Envelope detection techniques have applications in areas like medicine, sound classification and synthesis, seismology and speech recognition.
% Problem | Objectives
Nevertheless, a general approach to envelope detection doesn't exist, as most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation.
To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning.
% Methods | Procedure | Approach
The method draws inspiration from geometric concepts; the discrete signal is mapped to a set of points in the Cartesian plane, were a new measure of discrete curvature is used to identify the samples that are part of the envelope.
Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave.
% Results | Findings | Product
The algorithm compares favourably with classic envelope detection techniques based on smoothing, filtering and the Hilbert Transform, besides being physically plausible.
We provide visualizations of the envelope extracted for various real-world signals, with very diverse characteristics, such as voice, spoken and sang, and pitched and non-pitched musical instruments, and discuss some approaches to assess the quality of the obtained envelopes.
An optimized implementation is available via the Python Package Index (PyPI); interactive visualizations and source code are available at the accompanying repository.
\end{abstract}

{\bf Keywords:} alpha shapes; envelope detection; discrete curvature estimation; demodulation

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in both analogue and digital signal processing (DSP) \parencite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \parencite{2017LyonsDigital}, with most envelope detection techniques being designed to account for very specific settings, like pure sinusoids with moderate noise content, for example.

Those constraints exclude most physical signals, as is the case of recorded sound, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \parencite{2013Mengempirical}.

In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \textcite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of the intelligibility of Mandarin tones. This is also the case for the English language \parencite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.

The temporal envelope also helps to impart emotion and identity to the human voice \parencite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \parencite{2011LokkiEngaging}, for example.

Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting a priori knowledge of the waves, are prevalent, such as the one presented by \textcite{2014YangFast} for the distributed monitoring of fibre optic or the one formulated by \textcite{2018AssefModeling} in the context of medical ultrasound imaging.

The problem of envelope detection of a discrete signal, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.

Considering the standard 2d representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of an individual sample as a point. Provided a coherent mapping from this space to the standard cartesian plane is available, the original DSP problem can be transformed into a geometric one.

% TODO

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is using the concept of a convex hull, as defined in \textcite{2008BergComputational}. The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.

When the general shape of a set of points present prominent concavities, the convex hull overestimates the boundaries. To address this limitation \textcite{1983Edelsbrunnershape} introduced the concept of alpha shapes, that became known as the concave hull: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that only points that can be touched by a circle of radius $ \alpha $ coming from an infinite distance towards the baricenter of the set of points are part of the frontier. The polygon obtained linking those points forms the concave hull of this set.

Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property meaning that, for a given set of points and a given radius, only one concave hull exists. When the radius of the circle $ \alpha \to \infty $, concave hulls tend to the convex hulls.

Consider, for example, the artificially generated discrete wave in Fig. \ref{fig:Hulls} below, where the links between consecutive samples were deemphasized to highlight the "set of points" nature of a signal: The convex hull and a concave hull, in the form of an alpha shape with a particular value of $ \alpha $ are illustrated.

\begin{figure}[H]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, and the convex and concave hull at $ \alpha=300 $ defined by it.}
  \label{fig:Hulls}
\end{figure}

The alpha shapes approach is used in areas such as detection of features in images \parencite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \parencite{2015WuAutomated} and spectroscopy \parencite{2019XuModeling}.

This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, being particularly illustrative of the potential synergy between geometric and DSP approaches.

Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \textcite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \parencite{2015YangRepresenting}.

Following this path, we present a general approach to envelope detection, exploiting the intrinsic characteristics of a spectrally complex wave, in order to completely avoid the need for manual intervention or parameter tuning.

% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.

We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.

The rest of this paper is divided as follows: We start by defining as specifically as possible what a temporal envelope is, especially in the context of this work and providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.

We proceed to illustrate the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present in digital representations of waves.

Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave, via the procedure described in sequence.

In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature. We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, that can be seen as partial envelopes considering only the positive and negative pulses of a wave, respectively.

We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.

Interactive visualizations are available from the repository prepared for this work \parencite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available. Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Temporal envelope}\label{sec:Characterization of a temporal envelope}

The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves, often with well-defined frequencies \parencite{2011TurnerDemodulation}.

While there is no general formal definition of the envelope of broadband signals \parencite{2014Yangnovel,2015Yangtheoretical,2019Jiaempiricalj}, the general intuition is that an envelope is a smooth curve situated as close as possible to the signal \parencite{2019Jiaempiricalj}.

In this context, the most mathematically sound definition of an envelope involves the representation of its underlying wave as an analytic signal, as introduced by Gabor in 1946 \parencite{2007HahnHistory}. In his work, \textcite{1946GaborTheory} applies the then relatively new mathematical machinery of the quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.

This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \parencite{2016HePraat} where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $.

Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \parencite{2012DauSpeech, 2019Jiaempiricalj}, being even physically paradoxical in some cases \parencite{1996Loughlinamplitude, 2014Yangnovel}.

Fig. \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from the Fig. \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.

\begin{figure}[H]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half, the sinusoid is free of noise, while in the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

For a real world signal, such as an alto singer sustaining a steady note as shown in Fig. \ref{fig:AnalyticSignalAlto} the Hilbert transform alone provides a poor envelope, since it retains most part of the frequencies of the underlying wave. With further processing, however, such as filtering, an output closer to the intuitive definition of envelope given in the beginning of this section can be obtained.

\begin{figure}[H]
  \centering
    \includegraphics{images/01AnalyticSignalAlto.pdf}
  \caption{Representation of an alto singer sustaining a steady note and a comparison of the envelopes obtained by the Hilbert transform with and without further filtering.}
  \label{fig:AnalyticSignalAlto}
\end{figure}

In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \parencite{2011TurnerDemodulation}.

$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast varying part, called throughout the literature as its (temporal) fine structure, carrier component, or frequency modulation (FM).

In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same $ W(t) $ \parencite{2011TurnerDemodulation, 2013Mengempirical, 1996Loughlinamplitude}.
% TODO still ill posed
We address this problem by assuming $ C(t) $ normalized between $ \{-1, 1\}$. This assumption doesn't cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.

In this work we are concerned with the discrete version of this problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, instead of the continuous function $ W(t) $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.

The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals. In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the digital signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second. We can thus define:

\begin{subequations}\label{eq:Envelope}
\begin{align}
  \text{Given:}& \\
  \mathbf{w} & \in \mathbb{R}^n, \\
  \mathbf{w} & = (w_0, w_1, \cdots, w_{n-1}) \\
  \text{We define:}& \\
  \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
  \mathbf{e} & = (e_0, e_1, \cdots, e_{n-1}), w_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
  \mathbf{c} & = (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1 \\
  \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{align}
\end{subequations}

The $ \odot $ operator in \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors. Fig. \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by multiplying the normalized carrier wave \textbf{c} using the envelope \textbf{e}.

\begin{figure}[H]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of a previously known envelope \textbf{e} and carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

In Fig. \ref{fig:Envelope} a signal arising from the multiplication of a known carrier wave \textbf{c} and a known envelope \textbf{e} is shown; this helps to illustrate a key concept: note from Fig. \ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being in fact equal to the value of the wave $ w_i $. $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

% TODO

This realization that the envelope of a wave is defined only at some of the wave's local extrema is crucial, as we can use it to generate a simplified representation of a signal that will make the algorithm more efficient by considering only the local extrema of a given signal.

\section{Simplified Representation}

We start with the definition of a pulse: each time the value of $ w_i$ changes sign, that is, every time the amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This definition is in line with the one presented in \textcite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From \ref{fig:Envelope} and the discussion in Section \ref{sec:Characterization of a temporal envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to great computational economy.

For this reason we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_j \lvert) $, that is, $ \lvert w_j \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.

This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
While the exact relation between $ n $, the size of the original signal, and $ m $, the size of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals. In practice, $ m \ll n $ for most real world discrete signals.

\begin{figure}[H]
  \centering
    \includegraphics{images/03Pulses.pdf}
  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle. P is the set of the points $ P_j = (i_j, \lvert w_j \lvert) $ representing the absolute value of those extrema.}
  \label{fig:Pulses}
\end{figure}

\section{Mapping to Cartesian coordinates}

The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous amplitude of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.

Geometrically, this coordinate system has a basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in finding a new basis vector that would make the geometric representation consistent, in the sense that pulse's amplitude $ \lvert w_j \lvert $, now mapped to the ordinates of a point, and a pulse's length, approximately equivalent to the horizontal distance between two consecutive points in the geometrical interpretation would define, on average, a square.

To achieve that, we divide the basis vector of the original vertical axis by the average of all ordinates, effectively cancelling the unit of the vertical components of the points. We could divide the basis vector of the horizontal axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior, likewise eliminating the axis' unit.

The drawback is that we would lose the direct link between P and \textbf{w} that arises from the fact that the abscissas of the points in P are indices $ i $ of \textbf{w}: we choose instead to leave the abscissa intact and multiply the now unitless vertical basis axis by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior. In this way, both axes will have unit $ i $, and the relationship is preserved, making it easier to recover the envelope points later.

In this new coordinate system we have a basis vector $ \mathbf{b}^\prime = \{ \mathbf{b}^\prime_x, \mathbf{b}^\prime_y \}$ where $ \mathbf{b}^\prime_x = (i, 0) $ and $\mathbf{b}^\prime_y = (0, i) $, related to the original basis vector by the vertical scaling factor $ \beta $ as formalized in Eq. \ref{eq:Basis}:

\begin{subequations}  \label{eq:Basis} % TODO
\begin{align}
\mathbf{b}^\prime_x &= \mathbf{b}_x  \\
\beta &= \frac{\left( \frac{i_{m-1} - i_0}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
\mathbf{b}^\prime_y &= \beta \ \mathbf{b}_y 
\end{align}
\end{subequations}

The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Coordinates}, shown in scale for both coordinate systems.

\begin{figure}[H]
  \centering
    \includegraphics{images/04Coordinates.pdf}
  \caption{The set of points P and the set V of vectors between two adjacent points of P, in both the original and normalized coordinate systems, shown in scale.}
  \label{fig:Coordinates}
\end{figure}

\section{Discrete Curvature Estimation}

Having established a map from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system and, thus, a way to consistently transform our simplified description of a discrete signal into a set of points in the plane, the shape of which can be estimated with the use of geometric algorithms.

The goal of those algorithms, such as the convex hull and the alpha-shapes, is to discriminate which points of a set are part of its frontier.

Translating the explanation of the alpha-shapes algorithm presented in \textcite{1994EdelsbrunnerThree} to our context, we can say that the points that are part of the envelope are those that can be touched by a circle of a given radius $ r $, that must not contain any point of the signal, coming from $ y \to \infty $ towards the Cartesian representation of the signal.

% TODO formal definition of alpha shapes, maybe image

Intuitively, one can picture a circle of fixed radius being rolled above the signal, and marking the points it touches as envelope points.

To employ an alpha shapes inspired algorithm in this set of points, however, the radius $ r $ of such circle must be estimated, and for that a measure of discrete curvature is needed.

Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists.

The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \parencite{2001CoeurjollyDiscrete}.

Concerning direct methods \textcite{2014CarrollSurvey}, for example, derive three such definitions based on the approximation of a circle by an inscribed, centered and circumscribed polygon. In the context of three-dimensional meshes, \textcite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.

Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case). Instead, we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points.

Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity.

Thus, the set V of the vectors from each point of P to the next, shown in the Fig. \ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values thus obtained to infer the average curvature of the whole discrete curve.

\begin{figure}[H]
  \centering
    \includegraphics{images/04Vectors.pdf}
  \caption{The set of points P and the set V of vectors between two adjacent points of P, in the normalized coordinate system.}
  \label{fig:Vectors}
\end{figure}

We formalize the set of vectors V in Eq. \ref{eq:Vectors} below.

\begin{equation} \label{eq:Vectors}
\text{V} = \{ v_0, v_1, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
\end{equation}

\subsection{The Equivalent Circle Approach}

We thus proceed to define a discrete curvature measure over the edges of a discrete curve. To that end we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve. This definition is convenient, since it is equivalent to that of the osculating circle \parencite{2016VasaMesh}.

One of those edges, a vector $ v_k \in \text{V} $, is singled out in Fig. \ref{fig:DiscreteCurvature}. $ v_k $ represents the net displacement of the envelope from the extrema $ P_{j} $ to $ P_{j+1} $. Since we dont yet know which points of P are part of the envelope, instead of assuming the trajectory before $ P_{j} $ passes through  
$ P_{j-1} $, it is more prudent to assume that it can have passed anywhere in the line segment from 
% TODO
That is because all we can assume from $ v_k $ is that, in this current region of the envelope, it is able to a vertical displacement of $ v_{k,y} $ in the same time as it performs a horizontal displacement of $ v_{k,x} $. In average, we can consider the envelope entering from the horizontal direction.

The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose tangent has the same change in direction, in the same horizontal distance, as the vector of interest, as shown in Fig. \ref{fig:DiscreteCurvature}. The average curvature of P will then be obtained by the average of all radii $ r_k $.

From Fig. \ref{fig:DiscreteCurvature} one can see that $ r_k = v_{k,x} / \sin(\theta_k) $ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of the horizontal line and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ using the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.

\begin{figure}[H]
  \centering
    \includegraphics{images/05DiscreteCurvature.pdf}
  \caption{The tangent unit vector of the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
  \label{fig:DiscreteCurvature}
\end{figure}

The radius of the circle equivalent to each $ v_k $ is defined in Eq. \ref{eq:Radius}. This equation has the desirable property that $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line. 
% It is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally far from each other.

\begin{equation} \label{eq:Radius}
  r_k = \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}}
\end{equation}

For calculation purposes, however, this can cause a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of the radius. In this way, the problem is avoided, since $ v_{k,y} $ is guaranteed to be greater than zero. The average of curvatures is then inverted in order to obtain the average radius, as illustrated in Eq. \ref{eq:Curvature}.

\begin{equation} \label{eq:Curvature}
r = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
\end{equation}

We need now to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope. Generally, alpha-shapes based algorithms resort to the construction of the Delaunay triangulation of the set of points, that is latter filtered to contain only the outer edges. We can adopt a more straightforward method, as our points have a sequential structure.

It is worth noting that, although that is the case throughout this work, the original signal doesn't need to be sampled at equal intervals.

Let $ x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty $ be the coordinates of the center of a circle with radius $ r $, that can be placed anywhere in the upper plane of the coordinate system;

Let $ P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c $ be three different points of the set P;

For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two different points of P, that is, $ \forall \quad P_b, P_c $ such that $ (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2 $, $ P_a $ will be part of the set $ E $ of the points belonging to the envelope if and only if none of those circles contain $ P_a $, or $ P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2 $. Summarizing this, we have Eq. \ref{eq:InEnvelope}:

\begin{subequations} \label{eq:InEnvelope}
\begin{align}
& P_a \in \text{E} \leftrightarrow (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 > r^2, \\
& \forall \ P_b, P_c \mid (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 = r^2, \\
& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \quad 0 \le y_o < +\infty
\end{align}
\end{subequations}

The algorithm follows directly from the definition in \ref{eq:InEnvelope} after noting that the first and last members of P will always be part of the envelope, since they are reachable by a circle of any radius $ r $. From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.

The procedure is formalized in the algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \textcite{2020TarjanoEnvelope} can easily serve as pseudocode.

\begin{algorithm}[H]
\caption{Retrieve Envelope} \label{FindEnvelope}
\begin{algorithmic}
\STATE \textbf{Given} $ \text{P} = \{P_0, P_1, \cdots, P_{m-1}\}, r $, \textbf{circle}, a function that returns the center of a circle of a given radius passing through two points and \textbf{distance}, a function that returns the Euclidean distance between two points,
\STATE \textbf{Let} $ \text{E} = \{ P_0 \}, \text{id1} = 0, \text{id2} = 1 $
\STATE \textbf{Define} $ P_c = \text{point}, \text{empty} = \text{boolean} $
\WHILE {$ \text{id2} < m $}
  \STATE $ P_c \leftarrow \textbf{circle}(r, P_\text{id1}, P_\text{id2}) $
  \STATE empty $ \leftarrow $ \TRUE 
  \FOR{$ i = \text{id2} + 1 $ \TO $ m $}
    \IF {$ \textbf{distance}(P_c, P_i) < r $}
      \STATE empty $ \leftarrow $ \FALSE
      \STATE id2 $ \leftarrow \text{id2} + 1 $
      \STATE \textbf{break}
    \ENDIF
  \ENDFOR
  \IF {empty}
    \STATE E $ \leftarrow $ E $ \cup \ \{ P_\text{id2} \} $
    \STATE id1 $ \leftarrow $ id2
    \STATE id2 $ \leftarrow \text{id2} + 1 $
  \ENDIF
\ENDWHILE
\end{algorithmic}
\end{algorithm}

The set E obtained via algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points. The most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.
% TODO
It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points.

\section{Quality of an envelope}\label{sec:quality}

The lack of a precise definition of what an envelope is makes the assessment of the quality of an envelope a difficult task, for which there are no agreed-upon objective metrics. Recalling Fig. \ref{fig:Envelope}, however, and the accompanying discussion, if one divides, element-wise, the original signal by the envelope, the carrier wave thus obtained is expected to be bounded between -1 and 1.

Figs. \ref{fig:spoken_voice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted.

From Fig. \ref{fig:spoken_voice} it can be seen that the algorithm performs well under the carrier metric. This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.

\begin{figure}[H]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
  \label{fig:spoken_voice}
\end{figure}

Those characteristics are in stark contrast with those of the sang voice in Fig. \ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown; there, the wave is almost periodic, with a stable waveform and envelope.

\begin{figure}[H]
  \centering
    \includegraphics{images/06Graph_alto.pdf}
  \caption{Signal, envelope and carrier for a record of an alto singer.}
  \label{fig:singer_voice}
\end{figure}

Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an electric guitar. The variations in frequency in the signal didn't affect the envelope extracted.

\begin{figure}[H]
  \centering
    \includegraphics{images/06Graph_bend.pdf}
  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
  \label{fig:Bend}
\end{figure}

In an effort to formalize the assessment of the quality of temporal envelopes, \textcite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope. We comment briefly below that the algorithm here presented satisfies the four conditions presented.

% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The first one states that, if a signal is bounded in magnitude, then its envelope should be, as well. Our algorithm complies with this requisite, as the envelope is composed of selected samples from the original signal itself.

The second statement is that if the signal has a finite frequency range, that frequency range must not be exceeded by the envelope. This is only directly applicable in the continuous case, where this condition was proposed.

Nevertheless, if we picture both the wave and the envelope as samples from continuous functions, is easy to see that this is the case for our algorithm; since the envelope is composed by a subset of the samples of the original digital signal, in the limit where all the samples of the wave are also part of the envelope, we have that the maximum frequency for both is the same.

In all other cases, the maximum frequency of the envelope will be smaller than that of the original wave. The range of frequencies of the envelope, then, will always lie between zero and the maximum frequency of the original wave.

The third condition states what, for a periodic signal, the envelope should be a straight line with the intercept equal to the amplitude of this signal. Since all the local maxima have the same amplitude in the case of a periodic wave, our envelope will indeed be a straight line, with the same amplitude as the original wave. An example of such behaviour can be observed in Fig. \ref{fig:Comparison_sinusoid}.

The fourth and last condition states that if a wave is multiplied by a constant, the envelope should also be multiplied by the same constant. That is also the case, again because the points that form the envelope are points of the original wave itself.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off pratical applicability in favour of mathematical formality.
In \textcite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves. Fig. \ref{fig:RefEnvelopeTheory} illustrates this concept, for a family of sinusoids that share the same frequency and amplitude, and whose phases vary from 0 to $ 2 \pi $.

\begin{figure}[H]
  \centering
    \includegraphics{images/09RefEnvelopeTheory.pdf}
  \caption{Envelope as the boundary of the region filled by a family of curves, for a family of sinusoids with the same frequency and amplitude, modulated by a polynomial.}
  \label{fig:RefEnvelopeTheory}
\end{figure}

% TODO
In practical applications, one seldom has access to a family of curve, instead being interested in obtaining the envelope of a single signal. Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis, in order to cover the time of one period of the signal.
Fig. \ref{fig:RefEnvelope} shows the reference envelope for the sound of a tom.

\begin{figure}[H]
  \centering
    \includegraphics{images/10RefEnvelope.pdf}
  \caption{Reference envelope obtained by shifting the original signal, for part of a sample of a tom.}
  \label{fig:RefEnvelope}
\end{figure}

Since envelope is not smooth, being composed of a series of horizontal segments, as can be seen in \ref{fig:RefEnvelope}, besides being computationally expensive to calculate. However, arising from a mathematically sound base, it can serve as a reference to infer the quality of more empirical methods.

\section{Comparison with traditional algorithms}

Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \parencite[e.g.,][]{2018AssefModeling}.

Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms. Specifically, we compare the present method with the following approaches:

\begin{enumerate}
\item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a discrete signal, can provide an approximation of the variation of its amplitude in time. We use the Savitzky-Golay algorithm \parencite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial.
\item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \parencite{butterworth1930theory} low pass filter. We use an order 2 filter with a cut-off frequency of 10 Hz.
\item Hilbert Transform: We apply the Hilbert transform to the original signal, filtering the absolute value of the complex result with a Butterworth filter with a cut-off frequency of 1/10 of the fundamental frequency of the original signal.
\end{enumerate}

In Fig. \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. 
% TODO
Note that none of the conventional algorithms complies with the third condition proposed by \textcite{1996Loughlinamplitude}, that states that the envelope of a periodic wave should be a straight line.

\begin{figure}[H]
  \centering
    \includegraphics{images/07Comparison_sinusoid.pdf}
  \caption{Comparison of envelope detection algorithms for a simple sinusoid}
  \label{fig:Comparison_sinusoid}
\end{figure}

For a recording of the key 33 of a grand piano, where parts of the sound are highly percussive and noisy, the Hilbert approach presents considerable oscillations, as can be seen in Fig. \ref{fig:Comparison_piano}.

\begin{figure}[H]
  \centering
    \includegraphics{images/07Comparison_piano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
  \label{fig:Comparison_piano}
\end{figure}

In Fig. \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are consistently undershooting the original wave.

\begin{figure}[H]
  \centering
    \includegraphics{images/07Comparison_soprano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
  \label{fig:Comparison_soprano}
\end{figure}

Besides, the traditional methods demanded careful choice of parameters. In the case of the Hilbert transform, post-processing in the form of filtering was applied, and an appropriate cut-off frequency had to be chosen. That is in contrast with our method, that organically tunes itself via the automatic choice of the radius of the circle representing the average discrete curvature of the signal.

Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of section \ref{sec:quality}

\begin{table}[H]
\centering
\begin{tabular}{lcccc}
\hline
\multicolumn{5}{c}{\textbf{Average error per frame}} \\
Signal & Present Work & Smoothing & Low pass & Hilbert \\ 
\hline
alto  & 0.001 & 0.337 & 0.342 & 0.163 \\
bend  & 0.000 & 0.029 & 0.030 & 0.008 \\
brass & 0.001 & 0.107 & 0.107 & 0.038 \\
nonperiodic & 0.000 & 0.211 & 0.158 & 0.000 \\
piano & 0.003 & 0.016 & 0.016 & 0.007 \\
sinusoid & 0.000 & 0.132 & 0.131 & 0.000 \\
soprano & 0.000 & 0.155 & 0.155 & 0.044 \\
spoken\_voice & 0.006 & 0.109 & 0.110 & 0.045 \\
tom   & 0.001 & 0.011 & 0.015 & 0.005 \\
\hline
Average & 0.001 & 0.123 & 0.118 & 0.035 \\
frac. of max. & 0.010 & 1.000 & 0.961 & 0.281 \\
\end{tabular}
\caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification. The discrete waves used were normalized between -1 and 1. The average is between the absolute value of those waves and the envelope divided by two.}
\label{table:Error_comparison}
\end{table}

We can see from the table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error 33\% smaller, on average, than the other methods. It is also worth noting that the errors of the traditional methods are very similar between them, in agreement with the visualizations.

The times taken for the algorithms to process each wave are shown in \ref{table:Time_comparison}. 
The Python source for the tests, ass well as the samples used, are available at the repository dedicated to this work, in the file "Comparison of Envelope Algorithms.py".

The methods were performed as described in the beginning of this section, using the implementations available at the Scipy library, under the Signal processing module. The presented algorithm, is on average, the second-fastest.

\begin{table}[H]
\centering
\begin{tabular}{lcccc}
\hline
\multicolumn{5}{c}{\textbf{Time in seconds}} \\
Signal  & Present Work & Smoothing & Low pass & Hilbert \\ 
\hline
alto  & 0.010 & 0.096 & 0.007 & 0.006 \\
bend  & 0.009 & 0.146 & 0.005 & 0.067 \\
brass & 0.015 & 0.253 & 0.005 & 0.035 \\
nonperiodic & 0.002 & 0.064 & 0.003 & 0.007 \\
piano & 0.022 & 0.316 & 0.009 & 0.030 \\
sinusoid & 0.003 & 0.069 & 0.003 & 0.006 \\
soprano & 0.111 & 0.574 & 0.013 & 0.293 \\
spoken\_voice & 0.003 & 0.062 & 0.003 & 0.017 \\
tom   & 0.002 & 0.070 & 0.001 & 0.004 \\
\hline
Average & 0.020 & 0.183 & 0.005 & 0.052 \\
frac. of max. & 0.108 & 1.000 & 0.030 & 0.283 \\
\end{tabular}
\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope detection.}
\label{table:Time_comparison}
\end{table}

\section{Superior and inferior envelopes}

In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different superior and inferior contours; in those cases, the algorithm here presented can be used independently in the positive and negative pulses defined by the original wave to define two “envelopes”, that we are going to call superior and inferior frontiers.

Fig. \ref{fig:FullFrontiers} illustrates the frontiers of six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal. All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios.

\begin{figure}[H]
  \centering
    \includegraphics{images/08FullFrontiers.pdf}
  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented. For each wave, the region highlighted in black is shown in detail besides the whole wave. For each wave, the horizontal axis is the sample number $ i $, while the vertical axis is the normalized amplitude. All 6 waves were recorded at 44100 fps.}
  \label{fig:FullFrontiers}
\end{figure}

The frontiers are satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust in relation to the number of samples and the frequencies of the waves. That is a peculiarity of our algorithm: the traditional algorithms are unable to generate two distinct envelopes for the same wave.

\section{Approximated location of pseudo-cycles}

% TODO

The temporal envelope adds complexity to the spectral representation of a wave \parencite{2019TarjanoNeuro}, and an accurate description of the envelope, while describing the evolution of the instantaneous amplitude of a signal in time, would also simplify further spectral analysis. Fig. \ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in \ref{fig:Envelope}. For the carrier, the power spectrum presents two nonzero values, while the power spectrum for the wave composed of the carrier modulated by the envelope is more complex.

\begin{figure}[H]
  \centering
    \includegraphics{images/11Fourier.pdf}
  \caption{Fourier power spectrum for the wave and carrier shown in \ref{fig:Envelope}.}
  \label{fig:Fourier}
\end{figure}

Moreover, the algorithm developed in this work naturally divides a signal into its pseudo-cycles, pinpointing them in the time-domain, thus providing the building blocks for the reconstruction of the fine structure of the wave.

\begin{figure}[H]
  \centering
    \includegraphics{images/09PCs.pdf}
  \caption{Section of the lines formed by the positions of the positive and negative frontiers of the digital wave illustrated in \ref{fig:singer_voice}, as extracted by the algorithm presented, and their average.}
  \label{fig:PCs}
\end{figure}

One can see in Fig. \ref{fig:PCs} a section of the plot of where the local maxima and minima occurs in the recording of an alto singer show in \ref{fig:singer_voice}. The average between the two is also shown. With this information, one can split a discrete wave into its pseudo-cycles, as is illustrated in Fig. \ref{fig:AvgPc}: It shows the superposition of this extracted pseudo-cycles, after being normalized by length, and their average. This same average is shown shifted, to exemplify that approximate $C^0$ and $C^1$ continuity is achieved.

\begin{figure}[H]
  \centering
    \includegraphics{images/10AvgPc.pdf}
  \caption{Pseudo-cycles of the wave illustrated in \ref{fig:singer_voice}, segmented and superposed, and their average. The average is also shown shifted by one average period, to illustrate the average continuity.}
  \label{fig:AvgPc}
\end{figure}

\section{Conclusion}
This work fills a gap identified in the theory of digital signal processing, where the lack of general procedures for the accurate identification of the temporal envelope of arbitrary waves poses an obstacle to the complete description and eventual manipulation of signals.

By presenting a general approach for envelope detection, and eliminating the need for parameter tuning, we expect to facilitate further research at the many areas that rely on envelope detection techniques, and improve already existing algorithm that have envelope detection as an intermediary step.

Illustrating the efficiency of an algorithm inspired by geometric aspects of a discrete signal on a wide range of real-world signals, we also hope to encourage research in this direction.

The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavours to motivate further research in this topic, that can ultimately give rise to a more mathematically sound envelope theory, especially concerning broadband waves; suggestions in this direction are also presented.

While relevant in its own accord, the procedure here presented approximately isolates the individual pseudo-cycles of a wave, pinpointing them in the time-domain.

Similarly, by decomposing a wave into its envelope and carrier, and simplifying the representation of such carrier in the frequency-domain, a rich set of advancements, especially in the areas of sound compression and synthesis in the time and frequency-domains, can be built upon this initial algorithm.

By illustrating how this information can be used to obtain the average waveform of a discrete wave, we hope to encourage research in this area, ultimately leading to alternative representations of discrete waves.

\section{References}

\printbibliography[heading=none]

\end{document}