### 2021-05-11 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

\documentclass[preprint]{elsarticle}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts} % \mathbb
\usepackage{amssymb}  % \therefore
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{lineno,hyperref}

\bibliographystyle{elsarticle-num}
\begin{document}
\begin{frontmatter}

\title{Envelope Estimation using Geometric Properties of a Discrete Real Signal}

\cortext[cor1]{Corresponding author}
\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

\author{Carlos Henrique Tarjano Santos\corref{cor1}}
\ead{carlostarjano@id.uff.br}

\author{Valdecy Pereira}

\begin{abstract} 
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Envelope detection techniques have applications in medicine, sound classification, sound synthesis, seismology, and speech recognition. Nevertheless, a general approach to envelope detection doesn't exist. Most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation. To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning. The method draws inspiration from geometric concepts: the discrete signal is mapped to a set of points in the Cartesian plane, where a new measure of discrete curvature is used to identify the samples that are part of the envelope. Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave. The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert Transform, besides being physically plausible. We provide visualizations of the envelope extracted for various real-world signals with diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes. An optimized implementation is available via the Python Package Index (PyPI), while interactive visualizations and source code are available from the accompanying GitHub repository.
\end{abstract}

\begin{keyword}
  DSP \sep alpha shapes \sep envelope detection \sep discrete curvature estimation \sep demodulation
\end{keyword}

\end{frontmatter}

\maketitle
\linenumbers

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in analog and digital signal processing (DSP) \cite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for particular settings involving narrowband signals with moderate noise content.

Those constraints exclude most physical signals, such recorded sounds, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}.

In many contexts, however, the temporal amplitude envelope of a signal plays a prominent role: according to \cite{2017QiRelative}, for example, the envelope is at least as important as the fine structure of a sound wave in the context of intelligibility of Mandarin tones. This is also the case for the English language \cite{1995ShannonSpeech}, where even envelopes modulating mostly noise are still capable of conveying meaning.

The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music in general: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}.

Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting prior knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fiber optic, or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.

The problem of envelope detection of a discrete signal, traditionally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.

Considering the standard graphic representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents amplitude, one can think of each sample as a point in a plane. Provided a coherent mapping from this space to the Cartesian coordinate system is available, the original DSP problem can be transformed into the geometric problem of finding the shape of this set of points.

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is by using the concept of a convex hull, as defined in \cite{2008BergComputational}. The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.

To get an intuition, one can think of the points in a set as nails in a wooden board, around which a rubber band is stretched. The polygon defined by the rubber band is the convex hull of this set.
Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.

When the general shape of a set of points presents prominent concavities, the convex hull overestimates the boundaries. To address this limitation, \cite{1983Edelsbrunnershape} introduced the concept of alpha shapes, also known as concave hulls: a mathematically well-defined generalization of the convex hull of a finite set of points, closely related to the Delaunay triangulation and Voronoi diagrams of those points.

The intuition behind the alpha shapes algorithm is that, given a set of points and a fixed radius $ \alpha $, if it is possible to construct a circle containing one or more points of the set in such a way that the disc defined by the circle does not contain any other point of the set, then these points are part of the frontier. The polygon obtained linking all such points forms the alpha shape, or concave hull, of this set.

Fig. \ref{fig:AlphaShape} illustrates an alpha shape, for $ \alpha=10 $, of a set of points. Note that two points of the set lie on the circle illustrated, but the open disc it defines contains no points of the set. An analogous construction is possible for every point at the frontier.

Conversely, any circle of radius $ \alpha=10 $ containing at least one of the interior points defines an open disc that also contains at least another point of the set P.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00AlphaShape.pdf}
  \caption{The alpha shape of a set of points for $ \alpha=10 $. A circle of radius 10, containing two points of the set, but defining a disc that doesn't contain any point of the set, is also shown.}
  \label{fig:AlphaShape}
\end{figure}

Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one alpha shape exists. When the radius $ \alpha $ of the circle tends to infinity, the alpha shape of the set tends to its convex hull.

Consider, for example, the artificially generated discrete wave in Fig. \ref{fig:Hulls}, where the links between consecutive samples were de-emphasized to highlight the “set of points” nature of a signal: two alpha shapes, with different values of $ \alpha $, and the convex hull are illustrated. As $ \alpha $ increases, the contour of the set of points becomes coarser, and closer to the convex hull.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/00ConvexHull.pdf}
  \caption{A discrete signal interpreted as a set of points in the Cartesian plane, two alpha shapes, and the convex hull defined by it.}
  \label{fig:Hulls}
\end{figure}

Alpha shapes are used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from clouds of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.

This last work, which involves the estimation and removal of the Blaze function (a kind of envelope) of an echelle spectrograph, is particularly illustrative of the potential synergy between geometric and DSP approaches.

Other steps in the direction of applying geometric algorithms to envelope detection were made by \cite{2015YangSkeleton}, with an algorithm based on the construction of a skeleton underlying the digital wave of interest, and by \cite{2015YangRepresenting}, via the direct translation of computer vision methods to the task of envelope detection.

Following this path, we present a general approach to envelope detection, exploiting geometric characteristics of a spectrally complex wave to completely avoid the need for parameter tuning.

We start with a brief background of envelope detection in Section \ref{sec:envelope} where we define, as precisely as possible, what constitutes a temporal envelope.

In Section \ref{sec:representation} we explore some simplifications to the task of envelope detection that arise from the envelope definition adopted in this work.

Section \ref{sec:mapping} explores the geometric characteristics of envelope detection and formalizes how this classic DSP problem can be transformed into the geometric problem of finding the shape of a set of points.

We next explore, in Section \ref{sec:curvature}, the concept of discrete curvature, used in Section \ref{sec:envelopedetection} to discriminate which samples of a discrete signal belong to the envelope.

Section \ref{sec:quality} discusses the difficulties in assessing the quality of an envelope, suggesting qualitative and quantitative approaches. It is also demonstrated that the method presented conforms to the four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}. We also compare the presented algorithm with traditional envelope detection methods in terms of accuracy and processing time.

Section \ref{sec:extensions} discusses potential applications of the algorithm, some of which might not be immediately obvious.

Links to interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, which also contains an implementation of the method in pure Python.

A Python module can be installed from the Python Package Index (PyPI) with the command “pip install signal-envelope”. Those working under Windows 64-bit can benefit from a specialized dynamic library implemented in C++, automatically used within this module. This library can be obtained separately via the repository's release section.

\section{Temporal envelopes}\label{sec:envelope}

The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves with well-defined frequencies \cite{2011TurnerDemodulation}.

While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the intuition is that the ideal envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}, without intersecting with it.

The most mathematically sound definition of an envelope, introduced by Gabor in 1946 \cite{2007HahnHistory}, involves representing the underlying wave as an analytic signal. In his work \cite{1946GaborTheory}, Gabor employs the then relatively new mathematical machinery of quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform can be applied to a real signal in order to create an equivalent complex signal.

This complex signal became known as the analytic signal, and has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \cite{2016HePraat}, where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $. The analytic signal representation is common in traditional AM-FM literature.

Despite its widespread use, and effectiveness in the context of narrowband signals, envelope detection techniques based on the Hilbert transform don't behave well for broadband signals \cite{2012DauSpeech, 2019Jiaempirical}, being even physically paradoxical in some cases \cite{1996Loughlinamplitude, 2014Yangnovel}.

Fig. \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from Fig. \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected on the envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, obtained by the Hilbert Transform approach. In the first half of the samples, the sinusoid is free of noise. In the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the signal.}
  \label{fig:AnalyticSignal}
\end{figure}

For real-world signals, such the one of an alto singer sustaining a steady note partially shown in Fig. \ref{fig:AnalyticSignalAlto}, the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave. With further processing, such as filtering, a smoother output can be obtained.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignalAlto.pdf}
  \caption{Comparison of the envelopes obtained by the Hilbert transform with and without further filtering, for a section of the representation of an alto singer sustaining a steady note.}
  \label{fig:AnalyticSignalAlto}
\end{figure}

In general, the problem of envelope detection can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two waves $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \cite{2011TurnerDemodulation}.

$ E(t) $ represents the slow varying part of the wave, also known as the (temporal) envelope, modulator component, or amplitude modulation (AM) of $ W(t) $, while $ C(t) $ models its fast-varying part, called throughout the literature as its (temporal) fine structure, carrier wave, or frequency modulation (FM).

In the case of broadband signals, the problem of envelope detection of an arbitrary wave is ill-posed, in the sense that an infinite number of pairs of $ E(t) $ and $ C(t) $ can result in the same $ W(t) $ \cite{2011TurnerDemodulation, 1996Loughlinamplitude}.

We address this problem by assuming $ C(t) $ to be normalized between $ \{-1, 1\}$ such that $ |\max(C(t))| = 1 $. This assumption does not cause any loss of generality, since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.

In this work we are concerned with the discrete version of the envelope detection problem: given a finite discrete wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, obtaining the temporal envelope $ \mathbf{e} \in \mathbb{R}^n $ of $ \mathbf{w} $.

The continuous definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals. In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the discrete signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second. We can thus define:

\begin{align} \label{eq:Envelope}
\begin{split}
  \text{Given:}& \\
  \mathbf{w} & \in \mathbb{R}^n, \\
  \mathbf{w} & = (w_0, w_1, \cdots, w_{n-1}) \\
  \text{We define:}& \\
  \mathbf{e}, \mathbf{c} & \in \mathbb{R}^n \\
  \mathbf{e} & \triangleq (e_0, e_1, \cdots, e_{n-1}), e_i \ge 0 \ \forall \ i, \ 0 \le i \le n-1 \\  
  \mathbf{c} & \triangleq (c_0, c_1, \cdots, c_{n-1}), |c_i| \le 1 \ \forall \ i, \ 0 \le i \le n-1, |\max(\mathbf{c})| =1 \\
  \text{Such as:}& \\
  \mathbf{w} & = \mathbf{e} \odot \mathbf{c} \ \therefore \ w_i = e_i c_i \ \forall \ i, \ 0 \le i \le n-1
\end{split}
\end{align}

The $ \odot $ operator in Equation \ref{eq:Envelope} stands for the Hadamard product and denotes element-wise multiplication of two vectors. Fig. \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by element-wise multiplication between the normalized carrier wave \textbf{c} and the envelope \textbf{e}.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} arising from the element-wise multiplication of an envelope \textbf{e} and an carrier \textbf{c}, both previously known. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

Fig. \ref{fig:Envelope} helps to illustrate a key concept: only at some local extrema the envelope can be inferred without prior knowledge of the underlying carrier.

TODO
% being, in fact, equal to the value of the instantaneous amplitude of the wave $ w_i $. That is, $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

This realization that the envelope of a wave is defined only at some of the wave's local extrema is essential, as it can be used to generate a simplified representation of a signal, making the algorithm more efficient.

Note that, in Fig. \ref{fig:Envelope}, the carrier, the original signal and even the envelope are representations of similar entities: signals in the same coordinate system. This is in line with \cite{2011CaetanoImproved,2011TurnerDemodulation,2017LyonsDigital}, but in contrast with some representations found in the literature of amplitude modulation of narrowband signals \cite[e.g.][]{1996Loughlinamplitude}, where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.

\section{Simplified Representation of a signal} \label{sec:representation}

We start with the definition of a pulse: each time the value of $ w_i $ changes sign, that is, every time the instantaneous amplitude of the discrete wave \textbf{w} crosses the horizontal axis, the beginning of a pulse is defined, with the next crossing defining its end. This interpretation of a pulse is in line with the one presented in \cite{national1996telecommunications}, where a pulse is defined as a rapid change in the amplitude of a signal, followed by a fast return to the baseline value; zero, in our case.

From Fig. \ref{fig:Envelope} and the discussion in Section \ref{sec:envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to a tremendous computational economy.

For this reason, we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extremum of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa. This is illustrated in Fig. \ref{fig:Pulses}.

While the exact relationship between $ n $, the number of samples of the original signal, and $ m $, the cardinality of the set P, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals. In practice, $ m \ll n $ is true for most real-world signals.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/03Pulses.pdf}
  \caption{Example of a discrete wave \textbf{w} divided into pulses, of which the extrema are highlighted with a circle. P is the set of the points $ P_j = (i_j, \lvert w_{i_j} \lvert) $ representing the absolute value of those extrema.}
  \label{fig:Pulses}
\end{figure}

\section{Mapping to the Cartesian coordinate system} \label{sec:mapping}

The points in P are not fit for geometric interpretation yet, since the abscissa and ordinate of their orthogonal coordinate system have different units: in the vertical axis, we have a unit related to the instantaneous amplitude of the wave, while in the horizontal axis we have the index at which each extremum occurs, ultimately a time-related unit.

Geometrically, we can think of these points as points in a vector space on a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (i, 0), \mathbf{b}_y = (0, a) $, where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in representing those points on the Cartesian coordinate system, and for that we need a transformation that places those points on the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space.

To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively canceling the unit of the vertical components of the points. If we were to proceed similarly concerning the horizontal components,
the direct link between P and \textbf{w} would be lost.

It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior point.

In this way, we can consider both axes as having unit $ i $. Dividing them by $ i $, we obtain the effect of representing the points on the standard basis. Numerically, the relationship is preserved, since the abscissas of the points $ P_{\mathbf{s}} $ are numerically equivalent to the indices $ i $ of \textbf{w}, making it straightforward to recover the envelope points on the original coordinates later.

Considering $ P_{\mathbf{b}} $, a point of the set P represented on the original coordinate system and $ P_{\mathbf{s}} $ the same point represented on the Cartesian coordinate system, we formalize this transformation in Equation \ref{eq:Basis}:

\begin{align}\label{eq:Basis}
\begin{split}
P & = (x, y) \\
\beta &= \frac{\left( \frac{i_{j+1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{b}} &= x \mathbf{b}_x + y \mathbf{b}_y = x (i, 0) + y (0, a) = (x i, y a) \\
P_{\mathbf{s}} &= x \mathbf{s}_x + \beta y \mathbf{s}_y = x (1, 0) + \beta y (0, 1) = (x, \beta y) \\
&\forall \  P \in \text{P} \\ 
\end{split}
\end{align}

The effect of this transformation can be seen in the points represented in Fig. \ref{fig:Coordinates}, shown in scale for both coordinate systems. We can consider our points residing in the Cartesian coordinate system after the transformation.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Coordinates.pdf}
  \caption{The set P of points, in both the original (top) and standard (bottom) coordinate systems, shown in scale.}
  \label{fig:Coordinates}
\end{figure}

\section{Discrete curvature estimation} \label{sec:curvature}

Having established a transformation from the original space of the two-dimensional representation of a discrete wave to the Cartesian coordinate system, we are now able to apply the geometric concept of the shape of a set of points.

Algorithms, such as the convex hull and the alpha shapes, solve this problem by discriminating which points of a set are part of its frontier.

A formal definition of alpha shapes is given at \cite{1983Edelsbrunnershape}. In the context of this work, however, gaining the intuition of the algorithm is more important, and thus the explanation presented in \cite{1994EdelsbrunnerThree} can be more insightful.

The explanation, translated to our context, invites one to think of the plane where the points of P are represented as a soft material, such as Styrofoam. The points themselves are made of a harder material, such as rock, and occupy a fixed position in this plane. The alpha shapes algorithm consists of carving this plane from the outside of those points, with a circular eraser of fixed radius $ \alpha $. At the end of the process, the remaining Styrofoam that couldn't be erased due to the impeding rocks is the alpha-shape of this set of points.

Alternatively, one can picture a circle of fixed radius being rolled above the points in the set P, marking the points it touches as part of the envelope.

To employ an alpha shapes inspired algorithm, however, the radius $ r $ of such a circle must be estimated, and for that, a measure of discrete curvature is needed.

Discrete curvature estimation is an important task in image processing \cite{2010FleischmannNovel} for which no default definition exists.

The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \cite{2001CoeurjollyDiscrete}. Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity; this is generally not the case for approaches involving curve fitting.

Concerning direct methods, \cite{2014CarrollSurvey} presents three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon. In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators, from a multivariate point of view.

Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case). Since we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points, a definition of discrete curvature over the edges of a discrete curve would be preferable.

Thus, the set V of the vectors from each point of P to the next, shown in the Fig. \ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values obtained to infer the average curvature of the discrete curve.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/04Vectors.pdf}
  \caption{The set P of points and the set V of vectors between two adjacent points of P, in the Cartesian coordinate system.}
  \label{fig:Vectors}
\end{figure}

We formalize the set V of vectors in Equation \ref{eq:Vectors}.

\begin{align} \label{eq:Vectors}
\text{V} = \{ v_0, v_1, \cdots, v_{m-2} \} \quad \text{where} \quad v_k = P_{j+1} - P_j \quad \forall \quad 0 \le k \le m-2
\end{align}

\subsection{The Equivalent Circle Approach}

We thus proceed to define a discrete curvature measure over the edges of a discrete curve. To that end, we are going to apply the definition of smooth curvature as the rate of change of the unit tangent to a curve. This definition is convenient, since it is equivalent to that of the osculating circle \cite{2016VasaMesh}, the value we are interested in finding.

One of those edges, in the form of the vector $ v_k \in \text{V} $, is singled out in Fig. \ref{fig:DiscreteCurvature}. The vector $ v_k $ represents the envelope's displacement from the extrema $ P_{j} = (x_j, y_j) $ to $ P_{j+1} = (x_{j+1}, y_{j+1}) $.

We assume that, near the vector $ v_k $, the envelope defines a vertical displacement $ v_{k,y} = y_{j+1} - y_j $ and a horizontal displacement $ v_{k,x} = x_{j+1} - x_j $. Since we don't yet know which points of P are part of the envelope, we can't be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.

Nevertheless, assuming it passes through $ P_{j} $, and is limited to the displacements just observed, the envelope might have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before $ v_k $ in Fig. \ref{fig:DiscreteCurvature}, before reaching $ P_{j} $.

Generalizing those observations for all vectors $ v_k \in \text{V} $, we can consider that, on average, the envelope insides horizontally at $ P_{j} $ and smoothly changes from the horizontal direction to the direction of the vector $ v_k $.

The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose unit tangent performs the same change in direction, in the same horizontal distance. The average curvature of P will be obtained by the average of all radii $ r_k $.

From Fig. \ref{fig:DiscreteCurvature} one can see that $ r_k = v_{k,x} / \sin(\theta_k) $, where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of the horizontal line and the slope $ m_k = v_{k,y} / v_{k,x} $, defined by $ v_k $, using the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/05DiscreteCurvature.pdf}
  \caption{The unit vector tangent to the circle changes from the horizontal direction in $ \hat{u}_0 $ to an inclination of $ \theta_k $ in $ \hat{u}_1 $, $ \theta_k $ being the angle that the vector $ v_k $ makes with the horizontal direction.}
  \label{fig:DiscreteCurvature}
\end{figure}

The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}. This equation has the desirable property $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.

The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally distant from each other.

\begin{align} \label{eq:Radius}
  r_k = \frac{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}}{ v_{k,y}}
\end{align}

For calculation purposes, however, the formulation in Equation \ref{eq:Radius} poses a problem, as a zero could make its way into the denominator of a fraction; in the actual algorithm, the quantity calculated for each vector is the curvature itself, that is, the inverse of each radius $ r_k $. In this way, the problem is avoided, since $ v_{k,x} $ is guaranteed to be greater than zero. The average of curvatures is then inverted in order to obtain the average radius $ \overline{r} $, as illustrated in Equation \ref{eq:Curvature}.

\begin{align} \label{eq:Curvature}
\overline{r} = \left( \frac{\sum\limits_{k=0}^{m-2} \left( \frac{ v_{k,y}}{v_{k,x} \sqrt{v_{k,x}^2 + v_{k,y}^2}} \right)}{m-2} \right)^{-1}
\end{align}

\section{Identifying the envelope} \label{sec:envelopedetection}

We need to construct an algorithm, using the radius just obtained, to identify the points that belong to the envelope. Generally, alpha shapes algorithms resort to the construction of the Delaunay triangulation of the set of points, which is later filtered to contain only the outer edges. We can adopt a more straightforward approach, since our points have a sequential structure.

Let $ x_o, y_o \in \mathbb{R}, \quad 0 \le y_o $ be the coordinates of the center of a circle with radius $ r $, that can be placed anywhere in the upper plane of the coordinate system, and $ P_a, P_b \in \text{P} $ be two points of the set P;

For all circles of center $ x_o, y_o $ and radius $ r $ that can be constructed to pass through any two points of P, that is, all circles such that $ (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 $, $ P_a $ and $ P_b $ will be part of the set $ E $ of the points belonging to the envelope if and only if exists one such circle defining a disc that does not contain any other point of P. Summarizing this, we have Equation \ref{eq:InEnvelope}:

\begin{align} \label{eq:InEnvelope}
\begin{split}
& (P_{a,x} - x_o)^2 + (P_{a,y} - y_o)^2 = (P_{b,x} - x_o)^2 + (P_{b,y} - y_o)^2 = r^2 \\
& P_a, P_b \in \text{E} \leftrightarrow (P_{c,x} - x_o)^2 + (P_{c,y} - y_o)^2 > r^2 \ \forall \ P_c \in \text{P} - \{P_a, P_b\} \\
& P_a, P_b, P_c \in \text{P}, \quad P_a \ne P_c, P_b \ne P_c \quad \text{and} \quad x_o, y_o \in \mathbb{R}, \ 0 \le y_o
\end{split}
\end{align}

The algorithm follows directly from the definition in Equation \ref{eq:InEnvelope} after noting that the first and last points of P will always be part of the envelope. Since $ P_a $ can be equal to $ P_b $, it is always possible to construct a circle containing those points, by placing the center of the circle as far as possible to the left or right of the first or last point, respectively. The discs defined by those circles can't contain any other point.

From $ P_0 $, our first pivot point, we construct circles passing through $ P_1, P_2, \cdots, P_d $ until a circle that doesn't contain any point in P is found; $ P_d $ then becomes the new pivot point and is included in E, and this process continues until $ P_{m-1} $ is reached.

The procedure is formalized in algorithm \ref{FindEnvelope}. We don't provide similar algorithmic descriptions of the preceding steps since the Python implementation available at \cite{2020TarjanoEnvelope} can easily serve as pseudocode.

\begin{algorithm}[ht!]
\caption{Retrieve Envelope} \label{FindEnvelope}
\begin{algorithmic}
\STATE \textbf{Given} $ \text{P} = \{P_0, P_1, \cdots, P_{m-1}\}, r $, \textbf{circle}, a function that returns the center of a circle of a given radius passing through two points and \textbf{distance}, a function that returns the Euclidean distance between two points,
\STATE \textbf{Let} $ \text{E} = \{ P_0 \}, \text{id1} = 0, \text{id2} = 1 $
\STATE \textbf{Define} $ P_c = \text{point}, \text{empty} = \text{boolean} $
\WHILE {$ \text{id2} < m $}
  \STATE $ P_c \leftarrow \textbf{circle}(r, P_\text{id1}, P_\text{id2}) $
  \STATE empty $ \leftarrow $ \TRUE 
  \FOR{$ i = \text{id2} + 1 $ \TO $ m $}
    \IF {$ \textbf{distance}(P_c, P_i) < r $}
      \STATE empty $ \leftarrow $ \FALSE
      \STATE id2 $ \leftarrow \text{id2} + 1 $
      \STATE \textbf{break}
    \ENDIF
  \ENDFOR
  \IF {empty}
    \STATE E $ \leftarrow $ E $ \cup \ \{ P_\text{id2} \} $
    \STATE id1 $ \leftarrow $ id2
    \STATE id2 $ \leftarrow \text{id2} + 1 $
  \ENDIF
\ENDWHILE
\end{algorithmic}
\end{algorithm}

The set E obtained by the algorithm \ref{FindEnvelope} is a subset of P, and thus also a set of points in the Euclidean space.

Recall from Section \ref{sec:mapping} that the vertical coordinates of these points are currently scaled, in relation to the original signal, by the transformation described in Equation \ref{eq:Basis}. We could invert this transformation, but since we were careful to maintain the direct link between the horizontal coordinates in both coordinate systems, we can simply consider the horizontal coordinates of P as the indices of the samples of the original discrete signal, and retrieve the envelope at each point directly.

After converting E to the original coordinate system, the most straightforward way to transform those points into the vector $ \textbf{e} \in \mathbb{R}^n $ is via linear interpolation, the approach used in this work.

It is worth noting, however, that many other procedures are available, both for the interpolation and the smooth approximation of points. Depending on the intended use of the envelope, k-curves interpolation \cite{2017Yank} is a specially interesting approach, since it guarantees that the maximum curvature occurs at the interpolated points.

\section{Quality of an envelope} \label{sec:quality}

The lack of a precise definition of what an envelope is makes the assessment of the quality of an envelope a difficult task, for which there are no agreed-upon objective metrics. Recalling Fig. \ref{fig:Envelope}, however, and the accompanying discussion, if one divides, element-wise, the original signal by the envelope, the carrier wave thus obtained is expected to be bounded between -1 and 1.

Figs. \ref{fig:spokenvoice}, \ref{fig:singer_voice} and \ref{fig:Bend} present discrete waves \textbf{w}, their extracted envelope \textbf{e}, and the inferred carrier \textbf{c}, obtained after dividing, element-wise, the original signal by its extracted envelope.

From Fig. \ref{fig:spokenvoice} it can be seen that the algorithm performs well under the carrier criteria. This is especially remarkable for the case of spoken voice, a notorious complex wave with fast changes in pitch and high noise content.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_spoken_voice.pdf}
  \caption{Signal, envelope and carrier for a record of male voice uttering the word “amazing”.}
  \label{fig:spokenvoice}
\end{figure}

Those characteristics are in stark contrast with those of the singing voice in Fig. \ref{fig:singer_voice}, where a recording of an alto singer sustaining a steady note is shown. There, the wave is almost periodic, with a stable waveform and envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_alto.pdf}
  \caption{Signal, envelope and carrier for a record of an alto singer.}
  \label{fig:singer_voice}
\end{figure}

Fig. \ref{fig:Bend} shows the envelope in the case of a bend performed in an electric guitar. The fast frequency variations in the signal didn't affect the envelope extracted.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/06Graph_bend.pdf}
  \caption{Signal, envelope and carrier for a record of a bend performed on an electric guitar.}
  \label{fig:Bend}
\end{figure}

\subsection{Theoretical guarantees}

In an effort to formalize the assessment of the quality of amplitude and frequency modulations of a signal, \cite{1996Loughlinamplitude} proposed some theoretical conditions necessary, but not sufficient, to ensure the physical plausibility of an envelope.

We comment below on how those conditions translate to the discrete setting, and show that the algorithm proposed satisfies all the four conditions presented.

The first one states that, if a signal is bounded in magnitude, its envelope should also be bounded. Our algorithm complies with this requirement, as the envelope is composed of selected samples from the original signal itself.

More formally, we have the set P of the absolute values of the local extrema of the original signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ that contains, by definition, the global extrema of $ \mathbf{w} $.
The set E of the points that belong to the envelope is a subset of P. We can thus write Equation \ref{eq:Req1}:

\begin{align} \label{eq:Req1}
\begin{split}
\text{P} &= \{ P_0, P_1, \cdots, P_{m-1} \} \quad \text{where} \quad P_j = (i_j, \lvert w_{i_j} \lvert) \quad \forall \quad 0 \le j \le m-1 \\
\text{E} &\subset \text{P} \implies \max (\text{E}) \le \max (\text{P}) = \max (|w_0|, |w_1|, \cdots, |w_{n-1}|)
\end{split}
\end{align}

If the amplitude of the original signal is bounded, that is $ \max (|w_0|, |w_1|, \cdots, |w_{n-1}|) = b \in \mathbb{R} $, not only the amplitude of the envelope is bounded, as required, but is guaranteed to have the same bound $ b $.

The second statement dictates that if the signal has a finite frequency range, that frequency range must not be exceeded by the carrier wave.

To see that this condition holds for the proposed method, consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $, where all the absolute values of its local extrema are greater than one.

This requirement doesn't imply loss of generality since, for an arbitrary signal, we can assure this condition by dividing all samples by the value of the minimum absolute local extrema.

The relation $ \min(\text{E}) \ge 1 $ thus holds, since $ \text{E} \subset \text{P}$.

Provided that one uses a method that preserves the original bounds when interpolating the points in E, such as linear interpolation, we can be sure that all the points in the envelope vector $ \mathbf{e} = (e_0, e_1, \cdots, e_{n-1}) $ are also greater than one.

Recalling Equation \ref{eq:Envelope}, where the relation between the original signal, the carrier and the envelope was defined as $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we can write the magnitude of an arbitrary frequency of the original signal ($ m_\mathbf{w}(f) $) and, likewise, the magnitude of an arbitrary frequency of the carrier ($ m_\mathbf{c}(f) $), in terms of the Discrete Fourier Transform (DFT) as in Equation \ref{eq:DFT}:

\begin{align} \label{eq:DFT}
\begin{split}
m_\mathbf{w}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( w_k \cos(2 \pi k f / n) \right)^2 + \left( w_k \sin(2 \pi k f / n) \right)^2} \\
m_\mathbf{c}(f) &= \sum_{k=0}^{n-1} \sqrt{\left( \frac{w_k}{e_k} \cos(2 \pi k f / n) \right)^2 + \left( \frac{w_k}{e_k} \sin(2 \pi k f / n) \right)^2} 
\end{split}
\end{align}

For any given frequency $ f $, both $ m_\mathbf{w}(f) $ and $ m_\mathbf{c}(f) $ are nonnegative. Also, since $ e_k \ge 1 \ \forall \ k, 0 \le k \le n - 1 $ we establish that $ 0 \le m_\mathbf{c}(f) \le m_\mathbf{w}(f) \ \forall \ f $ which is a more restrictive condition than the original: If $ \mathbf{w} $ has a finite frequency range, the magnitude of frequencies outside that range are zero, as are those of the same frequencies of the carrier $ \mathbf{c} $.

The third condition states that a physically plausible algorithm should, for a signal $ \mathbf{w} $ of constant amplitude $ a > 0, \ a \in \mathbb{R} $ and constant frequency, of the form $ a \cos(\omega i + \phi) $, generate a constant envelope $ \mathbf{e} $ of magnitude $ a $ and a carrier wave $ \mathbf{c} $ of the form $ \cos(\omega i + \phi) $.

An example of such behaviour can be observed in Fig. \ref{fig:Comparison_sinusoid}, where the envelope of a simple sinusoid is shown. To see that this is always the case, note that all the extrema of a sinusoid are of the form $ a (\pm 1) $ and $ P_j = (i_j, \lvert a (\pm 1) \lvert) = (i_j, a) \ \forall P_j \in \text{P} $. As $ \text{E} \subset \text{P} $, we have that the envelope $ \mathbf{e} = (a, a, \cdots, a) $ and is, thus, constant. Likewise, since $ \mathbf{w} = \mathbf{e} \odot \mathbf{c} $, we have $ c_i = a \cos(\omega i + \phi) / a = \cos(\omega i + \phi) \ \forall \ i, 0 \le i \le n - 1 $.

The fourth and last condition states that if a wave is multiplied by a constant, its envelope should also be multiplied by the same constant.

Recall from Section \ref{sec:mapping} that, before the extraction of the envelope, the original points undergo a transformation to lie in the Cartesian coordinate system. After the extraction, the inverse operation is performed on the points of the Envelope.

Consider a discrete signal $ \mathbf{w} = (w_0, w_1, \cdots, w_{n-1}) $ and the set P of its absolute extrema, represented in the standard basis, $ \text{P} = \{ P_0, P_1, \cdots, P_{m-1} \}, \quad P_j = (i_j, \beta \lvert w_{i_j} \lvert) $, where $ \beta $ is obtained from Equation \ref{eq:Basis}.

Given a constant $ \alpha > 0, \alpha \in \mathbb{R} $, we can define the scaled signal as $ \mathbf{w}^\prime = (\alpha w_0, \alpha w_1, \cdots, \alpha w_{n-1}) $. Observing, from Equation \ref{eq:Basis}, that $ \alpha $ can be factored out of denominator of the summation, we have $ \beta^\prime = \beta / \alpha $ and the new set $ \text{P}^\prime = \{ P^\prime_0, P^\prime_1, \cdots, P^\prime_{m-1} \}, \quad P^\prime_j = (i_j, \frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ is equal to the set P for the signal before scaling.

This guarantees that the same subset of points E will be identified as part of the envelope in both cases.
In the respective original basis system, applying the inverse change of basis, those points are $ P_j = (i_j, \frac{\beta}{\beta} \lvert w_{i_j} \lvert)  \quad \forall \  P \in \text{E} $, for the original signal, and $ P^\prime_j = (i_j, \frac{\alpha}{\beta}\frac{\beta}{\alpha} \lvert \alpha w_{i_j} \lvert) $ for the scaled signal.

\subsection{Reference envelope}

In order to obtain a numeric indicator of the quality of an envelope, we construct a reference envelope, that trades off applicability in favor of mathematical formality.
In \cite{1992BruceCurves} the envelope is defined as the boundary of the region filled by a family of curves. Fig. \ref{fig:RefEnvelopeTheory} illustrates this concept, for a family of sinusoids that share the same frequency and amplitude, and whose phases vary from 0 to $ 2 \pi $.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/09RefEnvelopeTheory.pdf}
  \caption{Envelope as the boundary of the region filled by a family of curves, for a family of sinusoids with the same frequency and amplitude, modulated by a polynomial.}
  \label{fig:RefEnvelopeTheory}
\end{figure}

In practical applications, one seldom has access to a family of curves, instead being interested in obtaining the envelope of a single signal. Nevertheless, an approximation can be obtained, by shifting the original signal along its initial position in the horizontal axis by half of its period, and taking the absolute maximum value at each index $i$.
Fig. \ref{fig:RefEnvelope} shows the reference envelope for the sound of a tom.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/10RefEnvelope.pdf}
  \caption{Reference envelope obtained by shifting the original signal, for part of a sample of a tom.}
  \label{fig:RefEnvelope}
\end{figure}

The envelope obtained is not smooth, being composed of a series of horizontal segments, as can be seen in Fig. \ref{fig:RefEnvelope}, besides being computationally expensive to calculate. However, arising from a mathematically sound background, it can serve as a reference to infer the quality of empirical methods.

\subsection{Comparison with traditional algorithms}\label{sec:comparison}

Direct comparison with many of the more recent algorithms is made difficult by the unavailability of digital implementations of such works, many designed to process analog signals \cite[e.g.,][]{2018AssefModeling}.

Nevertheless, insight can be gained from a comparison of the results of the method here proposed with some of the most common envelope extraction algorithms. Specifically, we compare the present method with the following approaches:

\begin{enumerate}
\item Smoothing: This is a relatively simple procedure that, applied to the absolute values of a discrete signal, can provide an approximation of the variation of its amplitude in time. We use the Savitzky-Golay algorithm \cite{1964SavitzkySmoothing} with a window width of 3001 samples and a cubic polynomial fitting.
\item Filtering: The absolute value of the wave is filtered using a digital implementation of the Butterworth \cite{butterworth1930theory} low pass filter. We use an order 2 filter with a cut-off frequency of 10 Hz.
\item Hilbert Transform: We apply the Hilbert transform to the original signal, filtering the absolute value of the complex result with a Butterworth filter with a cut-off frequency of 1/10 of the fundamental frequency of the original signal.
\end{enumerate}

In Fig. \ref{fig:Comparison_sinusoid} we illustrate the envelope extracted by the conventional algorithms, contrasting it with the result of the method presented in this work, for a simple sinusoid. 
Only the present algorithm and the one based on the Hilbert transform comply with the third condition proposed by \cite{1996Loughlinamplitude}, which states that the envelope of a periodic wave should be a straight line.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_sinusoid.pdf}
  \caption{Comparison of envelope detection algorithms for a simple sinusoid. The envelope for the Hilbert transform and for the present work are coincident in the image.}
  \label{fig:Comparison_sinusoid}
\end{figure}

For a recording of the key 33 of a grand piano, where parts of the sound, specially at the start, are highly percussive and noisy, the Hilbert approach presents a considerable undershooting, as can be seen in Fig. \ref{fig:Comparison_piano}.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_piano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the key 33 of a grand piano}
  \label{fig:Comparison_piano}
\end{figure}

In Fig. \ref{fig:Comparison_soprano} all envelopes exhibit a similar movement along the samples, but the ones generated by traditional algorithms are consistently undershooting the original wave.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/07Comparison_soprano.pdf}
  \caption{Comparison of envelope detection algorithms for the recording of the vocalization of a soprano singer}
  \label{fig:Comparison_soprano}
\end{figure}

Besides, the traditional methods demanded careful choice of parameters. In the case of the Hilbert transform, post-processing in the form of filtering was applied, and an appropriate cut-off frequency had to be chosen. That is in contrast with our method, which organically tunes itself via the automatic choice of the circle's radius representing the average discrete curvature of the signal.

Table \ref{table:Error_comparison} presents the average of the absolute errors of the techniques, when compared with the reference envelope introduced in the end of Section \ref{sec:quality}

\begin{table}[ht!]
\centering
\begin{tabular}{lcccc}
\hline
\multicolumn{5}{c}{\textbf{Average error per frame}} \\
Signal & Present Work & Smoothing & Low pass & Hilbert \\ 
\hline
alto  & 0.001 & 0.337 & 0.342 & 0.163 \\
bend  & 0.000 & 0.029 & 0.030 & 0.008 \\
brass & 0.001 & 0.107 & 0.107 & 0.038 \\
nonperiodic & 0.000 & 0.211 & 0.158 & 0.000 \\
piano & 0.003 & 0.016 & 0.016 & 0.007 \\
sinusoid & 0.000 & 0.132 & 0.131 & 0.000 \\
soprano & 0.000 & 0.155 & 0.155 & 0.044 \\
spoken\_voice & 0.006 & 0.109 & 0.110 & 0.045 \\
tom   & 0.001 & 0.011 & 0.015 & 0.005 \\
\hline
Average & 0.001 & 0.123 & 0.118 & 0.035 \\
frac. of max. & 0.010 & 1.000 & 0.961 & 0.281 \\
\end{tabular}
\caption{Comparison of the average errors of the algorithm presented in this work with the most common methods of digital envelope identification. The discrete waves used were normalized between -1 and 1.}
\label{table:Error_comparison}
\end{table}

We can see from Table \ref{table:Error_comparison} that the method presented here consistently outperforms the traditional approaches for all waves tested, yielding an error of less than 2\% on average, of the error of the traditional methods.

The times taken for the algorithms to process each signal are shown in Table \ref{table:Time_comparison}. 
The Python source for the tests, as well as the samples used, are available at the repository dedicated to this work, in the file “Comparison of Envelope Algorithms.py”.

The methods were performed as described in the beginning of this section, using the implementations available at the Scipy library, under the Signal processing module. The presented algorithm, is on average, the second-fastest.

\begin{table}[ht!]
\centering
\begin{tabular}{lcccc}
\hline
\multicolumn{5}{c}{\textbf{Time in seconds}} \\
Signal  & Present Work & Smoothing & Low pass & Hilbert \\ 
\hline
alto  & 0.010 & 0.096 & 0.007 & 0.006 \\
bend  & 0.009 & 0.146 & 0.005 & 0.067 \\
brass & 0.015 & 0.253 & 0.005 & 0.035 \\
nonperiodic & 0.002 & 0.064 & 0.003 & 0.007 \\
piano & 0.022 & 0.316 & 0.009 & 0.030 \\
sinusoid & 0.003 & 0.069 & 0.003 & 0.006 \\
soprano & 0.111 & 0.574 & 0.013 & 0.293 \\
spoken\_voice & 0.003 & 0.062 & 0.003 & 0.017 \\
tom   & 0.002 & 0.070 & 0.001 & 0.004 \\
\hline
Average & 0.020 & 0.183 & 0.005 & 0.052 \\
frac. of max. & 0.108 & 1.000 & 0.030 & 0.283 \\
\end{tabular}
\caption{Comparison of the processing time of the algorithm presented in this work with the most common methods of digital envelope detection.}
\label{table:Time_comparison}
\end{table}

\section{Extensions} \label{sec:extensions}

In this section we comment briefly on some characteristics and extensions of the presented algorithm, in order to suggest possible applications beyond envelope extraction. One of the less obvious is, perhaps, its use in the segmentation of a signal into pseudo cycles.

\subsection{Superior and inferior envelopes}

In practice, is not uncommon for a discrete wave, especially in the case of sound, to present somewhat different superior and inferior contours; in those cases, the algorithm can be easily adapted to infer from the positive and negative pulses defined by the original wave to define two independent “envelopes”, that we are going to call superior and inferior frontiers.

If we split P, as defined in Section \ref{sec:representation}, into $\text{P}^+$, the set of the local maxima of the original signal and $\text{P}^-$, the set of its local minima, we can apply the remaining steps of the algorithm in those two sets independently, arriving at a positive and a negative frontier.

Fig. \ref{fig:FullFrontiers} illustrates these frontiers for six diverse discrete sound waves, as well as an in-detail view of the highlighted segment for each signal. All waves are records of physical sounds, chosen to represent the applicability of the algorithm in real-world scenarios. The implementation available at the PyPI repository provides an option to extract the frontiers.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/08FullFrontiers.pdf}
  \caption{Positive and negative frontiers of six digital waves, as extracted by the algorithm here presented. For each wave, the region highlighted in black is shown in detail besides the whole wave. For each wave, the horizontal axis is the sample number $ i $, while the vertical axis is the normalized amplitude. All 6 waves were recorded at 44100 fps.}
  \label{fig:FullFrontiers}
\end{figure}

The frontiers were satisfactorily detected in the case of harmonic and inharmonic sounds, and are robust about the number of samples and the frequencies of the waves. It is worth noting that the positive and negative frontiers are generally smoother than the unified envelope and seems to conform better to the contours of the underlying discrete signal.

\subsection{Simplified spectral representation}

The temporal envelope adds complexity to the spectral representation of a wave \cite{2019TarjanoNeuro}. An accurate representation of the envelope, while describing the instantaneous amplitude of a signal, would also simplify further spectral analysis.

Fig. \ref{fig:Fourier} shows the frequency-domain power spectrum for the wave and the carrier presented in Fig. \ref{fig:Envelope}. For the carrier, the power spectrum presents two nonzero values, while the power spectrum for the wave, composed of the carrier modulated by the envelope, is more complex.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/11Fourier.pdf}
  \caption{Fourier power spectrum for the wave and carrier shown in Fig. \ref{fig:Envelope}.}
  \label{fig:Fourier}
\end{figure}

Fig. \ref{fig:FourierBend} shows part of the power spectra for both the original wave and the carrier obtained by the present algorithm, for the guitar bend sound in Fig. \ref{fig:Bend}, around its fundamental frequency. Two prominent frequencies can be seen, corresponding to the initial and final notes of the bend. In the carrier, the other frequencies appear smoother.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/12Fourier.pdf}
  \caption{Fourier power spectrum for the wave and carrier of the sound of a guitar bend shown in Fig. \ref{fig:Bend}.}
  \label{fig:FourierBend}
\end{figure}

\subsection{Approximated location of pseudo-cycles}

The algorithm developed in this work naturally divides a signal into its pseudo cycles, pinpointing them in the time-domain, as illustrated in Fig. \ref{fig:singer_voice}, for the sound of an alto singer singing a sustained note, as in Fig. \ref{fig:singer_voice}.

The vertical lines mark the indices of the samples that belong to the positive frontier of the original signal, and the section between each vertical line isolates a pseudo cycle of the wave. Although two local maxima exist in each pseudo cycle, only the largest was marked by the algorithm.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/09PCs.pdf}
  \caption{Section of the digital wave illustrated in Fig. \ref{fig:singer_voice}, with vertical lines at the indices of the samples that belong to the positive frontier.}
  \label{fig:PCs}
\end{figure}

Therefore, using those indices, one can split a discrete wave into its pseudo cycles. One application of that is illustrated in Fig. \ref{fig:AvgPc}: It shows the superposition of the extracted pseudo cycles, after being normalized by length, and their average. They were also rotated to make their starting point closer to zero.

This same average is shown shifted, to exemplify that approximate $C^0$ and $C^1$ continuity is achieved. This can be seen as the average waveform of the original discrete signal, and can be used, along with the information of its envelope, to generate an alternative representation of the wave, useful, for example, for compression and classification purposes.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/10AvgPc.pdf}
  \caption{Pseudo-cycles of the wave illustrated in Fig. \ref{fig:singer_voice}, segmented and superposed, and their average. The average is also shown shifted by one average period, to illustrate the average continuity.}
  \label{fig:AvgPc}
\end{figure}

\section{Conclusion}
This work fills a gap identified in the theory of digital signal processing, where the lack of general procedures for the accurate identification of the temporal envelope of arbitrary waves poses an obstacle to the complete description and eventual manipulation of signals.

By presenting a general approach for envelope detection, and eliminating the need for parameter tuning, we expect to facilitate further research in the many areas that rely on envelope detection techniques, and improve already existing algorithms that have envelope detection as an intermediary step.

Illustrating the efficiency of an algorithm inspired by geometric aspects of a discrete signal on a wide range of real-world signals, we also hope to encourage research in this direction.

The present work also highlights some gaps in our current understanding and definition of temporal envelope and, in so doing, endeavors to motivate further research in this topic, that can ultimately give rise to a more mathematically rigorous envelope theory, especially concerning broadband waves; suggestions in this direction were also presented.

While relevant in its own accord, the procedure here presented approximately isolates the individual pseudo-cycles of a wave, pinpointing them in the time-domain.

Similarly, by decomposing a wave into its envelope and carrier, and simplifying the representation of such carrier in the frequency-domain, a rich set of advancements, especially in sound compression and synthesis, both in the time and frequency-domains, can be built upon this initial algorithm.

By illustrating how this information can be used to obtain the average waveform of a discrete wave, we hope to encourage research in this area, ultimately leading to alternative representations of discrete waves.

\bibliography{bibli}

\end{document}