### 2021-05-12 Papers/01_Pulses_Envelope/Response to Reviewers.tex
response to reviewers

\documentclass{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts} % \mathbb
% \usepackage{amssymb}  % \therefore
\usepackage{algorithm}
\usepackage{algorithmic}
% \usepackage{lineno,hyperref}

% \bibliographystyle{elsarticle-num}
\begin{document}

\title{Response to Reviewers}

\maketitle

\section{Reviewer 2}
This is an interesting research. It seems that the authors have already put this article publicly accessible at arXiv.org. The authors are suggested to consider and address the following issues:

\subsection{Writing needs to be refined.}
We extensively reviewed the text, rewriting parts of it, besides restructuring the sections to improve flow and coherence.

\subsection{Issues with the Highlights: maximum 85 characters, including spaces, per bullet point}
Highlights were rewritten to conform with the requirement of less than 85 characters.

\subsection{Abstract looks too long. Should be presented in a more concise way.}
We revised the abstract, removing superfluous information and verbose phrasing in order to make it more succinct, resulting in a reduction from the original 349 words to 248 words in the revised version.

\subsection{Citation of references should be in order and no jump of reference numbers.}
This issue was caused by the set of commands used to compile the TeX source, and we are sorry for the overlook. It was fixed in the revised version.

\subsection{What's the Abscissa in the highlights (50k, 100k, etc.) and some figures?}
The abscissas refer to the index $i$ of the discrete signal's samples. We revised the legends of figures 3 to 6, 10 to 17, 21 and 22 to include this information, and also made this more explicit in the paragraph where the discrete version of the problem is introduced (page 5, line 109).

\subsection{Legends in some figures are confusing and hard to distinguish, such as in Fig 2: signal and Hilbert Envelope.}
We reworked all figures, colour coding different entities when deemed helpful, with careful attention to the impact in the readability of the legends.

In the case of Fig. 2 in the original manuscript, now Fig. 3 in the revised version (page 4, line 90), we changed the colour of the Hilbert envelope to red, and the line indicating the start of the noisy section to black, making both stand apart from the signal.

\subsection{Figure 3: the illustration of the carrier c and wave w does not comply with the common sense of envelope and modulation?}
We changed the legend of the figure (page 6, Fig. 5, line 116, top of the page) and Equation 1 (page 5, line 110) to make more explicit that the example was constructed from a known carrier and envelope.

Besides, we added a paragraph in the revised version (page 6, line 123) calling attention to the different representations found in the literature.

% TODO
\subsection{Definition of the envelope is very vague and controversial. What's the relationship with the conventional amplitude modulation?}
We made more explicit that a general definition of envelope, especially in the case of broadband signals, is still an open question in the literature, adding recent citations to corroborate this.(page 4, line 76; page 5, line 100)

In an effort to better define an envelope, we added a general definition for the envelope of a family of curves (page 13, line 311), that motivated a rewrite of the metric used to assess the general quality of envelopes. This metric is now based on the construction of a reference envelope from an approximation of this definition to a discrete acenario (page 13, line 310).

Moreover, we greatly extended the discussion on how the proposed method satisfies the 4 conditions necessary for the plausibility of the amplitude and frequency modulation of a signal, proposed by \cite{1996Loughlinamplitude}, reinforcing the relationship with conventional amplitude modulation (page 12, line 261).

\subsection{For the proposed "Equivalent Circle Approach", an acceptable mathematical background and justification is missing.}
We expanded the explanation for the Equivalent circle, justifying the assumption of the horizontal initial direction, and making other assumptions explicit (page 9, line 190).

\subsection{It's quite doubtful about the assumption of $ w = e \odot c $?}
We revised the explanation about this equation to make clear that, besides being assumed in \cite{2011TurnerDemodulation}, the relation holds by definition in our work (page 5, equation 1, line 108).

% \cite{2019Jiaempirical}

We also noted that the definition based on the analytic signal is common in classic AM-FM literature that deals with narrowband signals.

\subsection{P8: the authors claimed that "the algorithm here presented satisfies the four conditions presented", however, no convincing evidence is provided. Please prove it mathematically.}
We added mathematical proofs for each one of the four conditions.

\subsection{How was the benchmarking carried out? Details on the implementation should be provided.}
We explained the benchmarking more extensively, citing that the implementations used were from the signal processing module of the Scipy Python library, and also made the source code for the tests, as well as the used samples, available at the repository dedicated for the work.

\section{Reviewer 4}
This paper proposes a new method for envelope estimation by using the geometric properties of a discrete real signal. Using discrete curvatures to determine the envelope of a discrete signal is an interesting idea and results obtained by this proposed method seem to be acceptable. However, the reviewer cannot recommend the publication of the paper in its current form. The authors need to address the reviewers below and revise the paper accordingly:

\subsection{The description of the key concepts is not clear. The paragraphs of the current paper were poorly written and readers cannot follow them easily. The authors are suggested to revise the paper extensively to improve its readability.}
We revised the text extensively, with the aim of improving the overall readability and making the flow of ideas more linear. To that end, the overall structure of sections and subsections was substantially changed.

We also extended the definition of the basic concepts, adding Figures to better illustrate

\subsection{The authors describe basic concepts of convex and concave hulls in an intuitive manner and show that “the idea is to identify the local extrema that tough the envelope”. But the reviewer cannot fully understand the logical connection between them. The authors are suggested to explain their idea in a more explicit manner.}

\subsection{The reviewer cannot understand why filtering is needed in the pre-processing of Hilbert transform. Clarify the reasoning and necessity.}
We explained in the revised text that the results of the pure Hilbert transform retain great part of the frequency content of the underlying wave, specially in the case of broadband signals, and added a figure to exemplify this effect in the case of a real world signal of an alto singer uttering a sustained note. We moved the filtering to the post-processing, however, and results for the Hilbert transform were improved substantially.

\subsection{Envelopes obtained by the proposed method seem to be good, but those obtained by other comparable methods, such as Hilbert transform, are not reasonable. The reviewer believes that results obtained by Hilbert transform should not be bad as shown by the authors.}
Besides moving the filtering to the post-processing phase, we changed the cut-off frequency of the filter, that was fixed at 100 Hz, to 1/10 of the fundamental frequency of the original signal, improving the envelopes obtained by the Hilbert transform substantially.

\subsubsection{Besides, its end effects are serious. Is it caused by the filtering, which is mentioned by the reviewer in comment 3?}
The effects were caused by filtering the wave prior to the transform. We changed the order of the operations, eliminating the effects in the revised version.

\subsubsection{Provide an explanation for the results in Figures 10, 11, and 12.}

\subsection{5. There are a lot of mistakes in the texts, please proofread the paper carefully. Some of them listed in the following:}

\subsubsection{a. Full name at the first time for the abbreviate: being particularly illustrative of the potential synergy between geometric and DSP approaches.}
We inserted the full name and abbreviation for DSP in its first use in the first paragraph of the revised version, and did the same with other abbreviations.

\subsubsection{b. Figure or table or equation should be illustrated:}

\begin{itemize}
  \item From 3 and the discussion in the preceding chapter;
  \item The algorithm follows directly from the definition in 5 after noting;
  \item In 10 we illustrate the envelope extracted by the conventional algorithms;
  \item The times taken for the algorithms compared here to process each wave are shown in 2;
  \item Figure 14 shows the frequency-domain power spectrum for the wave and the carrier presented in 3.
\end{itemize}

This issue arose from a misconception that those terms would be added during the latex compilation, discovered late in the process of elaborating the manuscript, that caused many of the faulty references to remained unnoticed. We took steps to ensure proper indication of figures, tables and algorithms in the revised version.

\subsubsection{c. Wrong equation, should be $ r_k = v_{k,x} / \sin(\theta_k) $ : From figure 6 is easy to see that $ r_k = v_{k,x}  \sin(\theta_k) $}
We thank the reviewer for pointing out this error. This mistake was fixed in the revised version. It turned out to be a typo, in the sense that the following equation, derived in part from this one, was correct. We seized the opportunity to also double check all other equations in the revised version.

\subsection{6. Figure 7 is shown in the manuscript, but there is no explanation on the main body for it. Please add an explanation in the revised paper.}
We added an explanation to the figure 7 in the revised paper, that illustrates the envelope of a guitar bend, as suggested.

\section{Reviewer 5}

\subsection{Once the carrier frequency is known the technique presented in this paper is a special kind of limited curvature interpolation technique.}

We apologize if we misunderstood the commentary, but we think we addressed it when making clear that the wave in Fig. XX 
was constructed from a previously known carrier wave and envelope; that is the only instance when the carrier frequency, being a sinusoid, was known.

\subsection{The author should improve the citation by referencing corresponding techniques, such as K-curve, and bounded curvature interpolation methods. It is interesting to investigate the application of such technique to signal envelope extraction.}
The K-curves algorithm involves an iterative optimization problem that would add a considerable amount of time to the interpolation step, since the number of interpolation points is usually in the hundreds. Our main concern, in the interpolation step, is not so much about curvature, but rather about boundedness between the interpolated values.

As for the curvature estimation step, we tested interpolation and fitting methods during the development of an estimate of the curvature of a discrete curve. Some computational problems arose during piecewise polynomial fitting for example, due to the general high dimensionality of the problem, and some artifacts with B-splines approach motivated us to pursue alternative, local methods, that culminated in the development of the equivalent circle approach. We plan an extension to the algorithm in the future, where the curvature will be allowed to vary along the time, and for that k-curves seems ideal, because of its guarantee of maximum curvature at the control points.

Moreover, we indicated in the revised version that k-curves might be an interesting option, depending on the requisites for the extracted envelope.

\bibliographystyle{unsrt}
\bibliography{bibli}

\end{document}